\section{Conclusion}
\label{sec:conclusion}

We ran the first controlled head-to-head ablation of six grounding mechanisms for LLM scientific idea generation, holding the backbone, topics, and output schema fixed and evaluating 245 ideas through an objective literature-distance metric, a 2-judge off-family rubric, and position-controlled pairwise comparisons. The answer to our research question is \emph{partial and conditional}: grounding does not uniformly produce better ideas; it \emph{steers} the novelty--feasibility trade-off. Deduction, literature comparison, and outcome imagination raise judged novelty and win pairwise against unguided generation---literature comparison most strongly ($79\%$), and it alone improves idea-set diversity---while small experiments, induction, and proposal writing raise feasibility at novelty's expense.

\para{Key takeaway.} No mechanism increases objective distance from prior work, and LLM-judged novelty is uncorrelated with that objective measure ($\rho = 0.01$) even though the two judges agree with each other. The hypothesized ``better novelty'' is real in human-style judgment but a mirage under a grounded metric. Concretely: for judged novelty, diversity, and head-to-head wins use literature comparison; for feasibility use induction or proposal writing; avoid small-experiment grounding when novelty matters, as it rediscovers known methods; and for evaluation, never report single-LLM-judge novelty alone---pair it with a literature-distance metric and pairwise comparison, because they can disagree completely.

\para{Future work.} Four directions follow. (1)~Replicate across backbones (Claude, Gemini, DeepSeek) to test generality. (2)~Combine mechanisms---\eg deduction followed by literature-comparison filtering---to seek novelty \emph{and} feasibility together. (3)~Build the missing item-level human anchor on fresh, post-cutoff ideas to calibrate which metric tracks expert judgment. (4)~Test true small-experiment grounding in a runnable prompting benchmark where ideas can be executed in-loop. We release our code, the 245 generated ideas, judge scores, and analysis to support these extensions.
